Tracing the Hercules stream around the Galaxy

### 2010-05-20 tex/hercules.tex
groups-simulations: small tweaks

\begin{abstract}
It has been proposed that the Hercules stream, a group of co-moving
stars in the Solar neighborhood offset from the bulk of the velocity
distribution, is the result of resonant interactions between stars in
the outer disk and the Galactic bar. Only seen so far in the immediate
Solar neighborhood, I predict the distribution of stellar velocities
and the run of the Hercules stream as a function of location in the
Galactic disk in a simple model for the Galaxy and the bar that
produces the observed Hercules stream. I further show that the
Hercules feature is strong enough that it can be unambiguously
detected in the distribution of line-of-sight velocities in selected
directions. I quantify the most promising lines-of-sight for detection
in radial velocities: The most promising directions to target are at
$l \gtrsim 250^{\circ}$. The predictions presented here are only
weakly affected by distance uncertainties, assumptions about the
distribution function of the old stellar disk, and the details of the
Galactic potential including the effect of spiral structure;
\Gaia\ and future spectroscopic surveys of the Galactic disk will be
able to robustly test the origin of the Hercules stream and constrain
the properties of the Galactic bar.
\end{abstract}

@@

The velocity distribution of nearby stars contains various large
overdensities of co-moving stars that cannot be explained as the
vestiges of the clumpy nature of star formation
\citep[\eg,][]{Bovy10a}. In particular, the Hercules moving group, a
group of stars significantly offset from the bulk of the observed
velocity distribution \citep[\eg,][see
  \figurename~\ref{fig:obs}]{Dehnen98b,Bovy10a}, displays a wide range
of ages and metallicities
\citep{Blaauw70a,raboud98a,caloi99a,Bensby07a}. A compelling model
that has it originating through resonant interactions of stars in the
outer disk with the bar in the central region of the Galaxy has been
put forward \citep{dehnen00a,fux01a}, but any confirmation of this
picture beyond the locallly observed kinematics of stars remains
lacking. An investigation of the Hercules members' color--magnitude
diagram found no evidence for any significant metallicity anomaly with
respect to other local stars \citep{Bovy10a}, even though such a
difference should be expected as the Hercules stars originate from a
few kiloparsec toward the Galactic center where the average
metallicity is higher than in the Solar neighborhood. The absence of
the predicted metallicity anomaly could potentially be explained
through, \eg, radial mixing \citep{sellwood02a}, but it is clear that
chemodynamical modeling of the Galactic disk is currently not up to
the task of separating out these effects.

@@

To simulate the effect of the bar on stellar orbits near the Solar
circle, we follow the approach of \citet{dehnen00a}. For a fixed
position in the Galaxy, we evaluate the velocity distribution function
at values for the radial---toward the Galactic center---and
tangential---in the direction of Galactic rotation---velocity by
backward-integrating these velocities in the the Galactic potential to
obtain the intial orbit before bar formation. We posit that at this
epoch the old stellar disk was in a steady-state that can be described
by a simple distribution function that is a function of the energy and
angular momentum of the orbit alone. It follows from the collisionless
Boltzmann equation that the value of the velocity distribution at the
present position and velocity is equal to that of the steady-state
distribution function before bar formation evaluated at the initial
orbit.

Since the Hercules stream is only apparent in the planar motions in
the disk \citep[\eg,][]{Bovy09a}, we only consider a two-dimensional
model for the Galaxy. We use a simple power-law rotation curve to
model the axisymmetric Galactic potential
\begin{equation}
v_c = \vo(R/\Ro)^\beta \,,
\end{equation}
and for most of the simulations below we specify this model further to
a flat rotation curve ($\beta = 0$). The model for the bar is given by
the following potential
\begin{equation}
\Phi_b(R,\phi) = \Ab(t) \cos [2(\phi - \Omegab t)] \times
\left\{ \begin{array}{ll} -(\Rb/R)^3\,, & \mathrm{for}\ R \geq
  \Rb\,,\\ (R/\Rb)^3-2\,, & \mathrm{for}\ R \leq \Rb\,. \end{array}
\right.
\end{equation}
Here \Omegab\ is the pattern speed of the bar and \Rb\ is the bar
radius, which is fixed to be 80\,percent of the bar's corotation
radius. The bar is grown smoothly during a time $t_1$, always set to
half of the total integration time in what follows, following the
prescription
\begin{equation}
\Ab(t) = \alpha \, \frac{\vo^2}{3}\,\left(\frac{\Ro}{\Rb}\right)^3
\times \left\{ \begin{array}{lll}
\frac{3}{16}\xi^5 - \frac{5}{8} \xi^3 + \frac{15}{16} \xi + \frac{1}{2}\,, & \xi \equiv 2\frac{t}{t_1}-1\,, & t \leq t_1\,,\\
1\,, & & t \geq t_1\,.\end{array} \right. 
\end{equation}
A comparison of the local velocity distribution obtained by using this
simple bar potential with that resulting from the more realistic bar
potential of \citet{gardner10a} shows that this simple model is
sufficient to model the bar's influence on stellar orbits in the outer
disk.

For the distribution function of the stellar disk before bar
formation, we use a Dehnen distribution function \citep{dehnen99b}
given by
\begin{equation}\label{eq:fdehnen}
\fdehnen(E,L) \propto \frac{\Sigma(\rE)}{\sigmaR^2(\rE)} \, \exp\left[ \frac{\Omega(\rE)\left[L-\Lc(E)\right]}{\sigmaR^2(\rE)}\right]\,,
\end{equation}
where \rE, \Lc, and $\Omega(\rE)$ are the radius, the angular
momentum, and the angular frequency, respectively, of the circular
orbit with energy $E$. Using the procedure given in \sectionname~3.2
of \citet{dehnen99b}, we choose the $\Sigma(R)$ and $\sigmaR$
functions such that they reproduce a disk with exponential surface
density and velocity dispersion profiles 
\begin{equation}
\Sigma(R) = \Sigma_0 \exp\left(-R/\Rs\right)\,, \qquad 
\sigmaR(R) = \sigma_0 \exp\left(-R/\Rsigma\right)\,,
\end{equation}
to an accuracy of a fraction of a percent at all radii. Our fiducial
model for the distribution function has \Rs = \Ro /3 and \Rsigma =
3\Rs, and $\sigma_0$ such that $\sigmaR(\Ro) = 0.2 \vo$.

As shown by \citet{dehnen99c,dehnen00a}, the model described above
that places the Sun at \Ro = 0.9\Rolr---where \Rolr\ is the radius at
which a circular orbit is in the outer Lindblad resonance with the
bar---at a current angle of 25$^{\circ}$ with the bar, with a bar
strength $\alpha = 0.01$, and integrating for four bar periods,
reproduces the Hercules feature in the local velocity distribution
while agreeing with other photometric and kinematical observations of
the bar \citep[\eg,][]{binney97a,bissantz02a,cole02a,Shen10a}. In what
follows, unless indicated otherwise, we fix all of the parameters of
the model at these fiducial values.

In what follows, all velocities are always considered to be with
respect to their local standard of rest in the absence of a bar.

@@

\citet{dehnen00a} only simulated the velocity distribution at the
position of the Sun for various combinations of the model parameters
defined in the previous Section. Nevertheless, his results on how the
velocity distribution at the Sun changes with the angle between the
Sun-Galactic center line and the bar could also be interpreted as
showing the velocity distribution on the Solar circle for different
Galactocentric azimuths, fixing the bar angle. In this way, these
simulations become predictions for what one might find when one
studies the velocity distribution in locations beyond those probed by
the \hipparcos\ mission.

In \figurename~\ref{fig:rphi2d}, I present the predictions for the
velocity distributions near the Solar circle for the full
360$^{\circ}$ range in Galactocentric azimuths for the fiducial model
from \sectionname~\ref{sec:method}. Only the range
$-90^{\circ} \leq \phi \leq 90^{\circ}$ is shown here, since the
barred Galaxy is symmetric with respect to rotations through
180$^{\circ}$, such that the predictions for the other side of the
Galaxy can be obtained from those shown by shifting the azimuths by
180$^{\circ}$. The $R = \Ro$ locus is not shown here, but it can be
inferred from \figurename~2 in \citet{dehnen00a}.

It is clear from this \figurename\ that we can trace the Hercules
stream going around the Galaxy and when looking further in or further
out. The effect of the bar is also not confined to a narrow range in
radii, rather the effects of the bar and especially its outer Lindblad
resonance are felt through a large range of radii (0.6 $\leq R/\Ro \leq
1.4$). These effects are strong and robust predictions of the
bar-origin model for the Hercules stream, especially in regions beyond
the Solar radius, where the Hercules feature can be found
significantly removed from the bulk of the velocity distribution,
which might be severely shaped at all locations by the processes that
create the pattern of low-velocity moving groups in the local velocity
distribution (the main mode in \figurename~\ref{fig:obs}). As I
discuss below, \Gaia\ will be able to easily verify these predictions.

@@

Rather than wait for a comprehensive survey of the disk of the Galaxy
that measures the full phase-space position of a large number of stars
by measuring proper motions, we can ask whether a targeted survey of
line-of-sight velocities in selected regions of the Galaxy could see a
clear signature of the Hercules feature. That this should be possible
follows from the following argument. The Hercules feature stands out
in the velocity distribution most in its Galactocentric tangential
velocity $\vphi$. On the Solar circle near the Sun we see almost
entirely this Galactocentric tangential velocity in the line-of-sight,
so we would expect to see at least a hint of the Hercules stream in
the line-of sight.

In \figurename~\ref{fig:rphi1d} I show the predicted line-of-sight
velocity distributions in various locations near the Solar circle all
around the Galaxy, obtained by integrating the full two-dimensional
velocity distribution over the tangential velocity. This
\figurename\ is no longer symmetric with respect to shifts of
180$^{\circ}$ degree in azimuth, since the line-of-sight direction
breaks this symmetry. The gray curve is the line-of-sight distribution
one would observe if the targeted stars were in a steady state in an
axisymmetric potential with a Dehnen distribution function
(\eqnname~[\ref{eq:fdehnen}]). Both line-of-sight velocity
distributions are normalized to have equal mass since I assume that
our knowledge of the Galaxy and the tracer population will not be
great enough that we can test predicted differences in the absolute
number of stars at a given location between the axisymmetric and the
bar-resonance predictions. These predictions are also largely a
function of the particular form and parameters of the distribution
function that we chose the represent the stellar disk before bar
formation.

We see that the prospects to see the Hercules feature are quite good
on our side of the Galaxy. The Hercules feature is especially strong
in the line-of-sight velocity distribution for Galactic longitudes $l
\geq 250^{\circ}$. This is because of the fortunate coincidence that
for those lines-of-sight we see velocities in the direction along
which the Hercules streams separates most from the bulk of the
velocity distribution (see \figurename~\ref{fig:rphi2d}) and that at
these azimuths the Hercules stream is at the furthest distance from
the rest of the velocity distribution. We see hardly any effect of the
bar on the line-of-sight velocity distributions on the other side of
the Galaxy ($90^{\circ} \leq \phi \leq 270^{\circ}$). This is due to
the fact that on that side we see mostly the Galactocentric radial
velocities. Even though upcoming surveys such as \apogee\ have the
ability to probe the other side of the Galaxy, for the purposes of
detecting the Hercules stream it is uninteresting.

To further analyze the best directions to target to best see the
Hercules feature in the line-of-sight velocities, I have calculated
the line-of-sight velocity distribution on a 100 by 20 grid in
Galactocentric azimuth and radius. As a measure of the difference
between the bar-resonance model and the steady-state, axisymmetric
model, I have calculated the Kullback-Leibler divergence between these
predicted distributions \citep[\eg,][]{mackay}
\begin{equation}
D_{\text{KL}}(P||Q) = \int \dd x\,p(x)\,\ln \frac{p(x)}{q(x)}\,,
\end{equation}
where, in this case, $x$ is the line-of-sight velocity, $p(x)$ is the
predicted distribution for the bar-origin model of the Hercules
stream, and $q(x)$ is the axisymmetric prediction. The
Kullback-Leibler divergence is only one among many possible choices to
quantify the difference between two probability distributions. It can
be interpreted as the expected weight of evidence for the prediction
by the bar-resonance model over the fully axisymmetric prediction that
can be expected from each sampled star. As such, locations with the
maximal Kullback-Leibler divergence are those in which the Hercules
feature stands out most prominent in the line-of-sight.

@@

Since the extinction in the optical bandpasses is rather large in the
directions identified as promising in \figurename~\ref{fig:detect}
\citep[$A_V \approx 8$ mag toward $l \approx
  270^{\circ}$;][]{marshall06a}, spectroscopic surveys in the
near-infrared such as \apogee\ are the best bet for seeing the
Hercules feature in the line-of-sight direction. The range of Galactic
longitudes that \apogee\ can target along the disk is $-5^{\circ} \leq
l \leq 250^{\circ}$ (SDSS-III Project Description\footnote{Available
  at \url{http://sdss3.org/collaboration/description.pdf}~.}). From
\figurename~\ref{fig:detect} this limit seems to be right on the edge
of where one might hope to detect the Hercules feature. Since
\apogee\ is expected to start taking data in 2011 and to finish
operations before the first \Gaia\ data release, \apogee\ could trace
the Hercules stream around the Galaxy \Gaia. Since \apogee\ will also
obtain detailed chemical abundances for its target stars, this would
also allow for a more detailed analysis of the abundances of stars in
the Hercules stream \citep[following, \eg,][]{Bensby07a,Bovy10a},
which could shed more light on radial mixing in the Galactic disk.

@@

\bibitem[Caloi \etal(1999)]{caloi99a}
  Caloi,~V.~, \etal, 1999,
  \aap, 351, 925

@@

\bibitem[Marshall \etal(2006)]{marshall06a}
  Marshall,~D.~J.~, \etal, 2006,
  \aap, 453, 635

@@

\clearpage
\begin{figure}
\includegraphics[width=0.5\textwidth]{veldist.eps}
\caption{The velocity distribution of nearby ($\lesssim 100$ pc) stars
  from \citet{Bovy09a}. \vR\ and \vphi\ are the velocities toward the
  Galactic center and in the direction of Galactic rotation,
  respectively. The Hercules moving group is visible as an overdensity
  at \vphi $\approx$ -0.2 $\vo$.}\label{fig:obs}
\end{figure}

\clearpage
\begin{figure}
\includegraphics[width=\textwidth]{rphi2d.ps}
\caption{Two-dimensional velocity distributions along the Solar
  circle. The axes of each subpanel are the radial and tangential
  velocities in Galactocentric cylindrical coordinates for the same
  ranges as in \figurename~\ref{fig:obs}. The velocity distributions
  on the other side of the Galaxy are obtained by shifting the
  azimuths by 180$^{\circ}$.}\label{fig:rphi2d}
\end{figure}

@@

\clearpage
\begin{figure}
\includegraphics[width=0.5\textwidth]{distuncertain.ps}
\includegraphics[width=0.5\textwidth]{df.ps}\\
\includegraphics[width=0.5\textwidth]{slope.ps}
\includegraphics[width=0.5\textwidth]{barstrength.ps}
\caption{Variation of the predicted velocity distribution at R = 1.075
  \Ro , $\phi = -21^{\circ}$ (d= 0.4 \Ro, $l$ = 270$^{\circ}$) with
  distance uncertainties, parameters and shape of the distribution
  function, slope of the rotation curve, and bar
  strength.}\label{fig:1dvar}
\end{figure}

>>>>

\begin{abstract}
It has been proposed that the Hercules stream, a group of co-moving
stars in the Solar neighborhood offset from the bulk of the velocity
distribution, is the result of resonant interactions between stars in
the outer disk and the Galactic bar. Only seen so far in the immediate
Solar neighborhood, I predict the distribution of stellar velocities
and the run of the Hercules stream as a function of location in the
Galactic disk in a simple model for the Galaxy and the bar that
produces the observed Hercules stream. I further show that the
Hercules feature is strong enough that it can be unambiguously
detected in the distribution of line-of-sight velocities in selected
directions. I quantify the most promising lines-of-sight for detection
in radial velocities; These are at $l \gtrsim 250^{\circ}$. The
predictions presented here are only weakly affected by distance
uncertainties, assumptions about the distribution function of the
stellar disk, and the details of the Galactic potential including the
effect of spiral structure; \Gaia\ and future spectroscopic surveys of
the Galactic disk will be able to robustly test the origin of the
Hercules stream and constrain the properties of the Galactic bar.
\end{abstract}

@@

The velocity distribution of nearby stars contains various large
overdensities of co-moving stars that cannot be explained as the
vestiges of the clumpy nature of star formation
\citep[\eg,][]{Bovy10a}. In particular, the Hercules moving group, a
group of stars significantly offset from the bulk of the observed
velocity distribution \citep[\eg,][see
  \figurename~\ref{fig:obs}]{Dehnen98b,Bovy10a}, displays a wide range
of ages and metallicities
\citep{Blaauw70a,raboud98a,caloi99a,Bensby07a}. A compelling model
that has it originating through resonant interactions of stars in the
outer disk with the bar in the central region of the Galaxy has been
put forward \citep{dehnen00a,fux01a}, but any confirmation of this
picture beyond the locallly observed stellar kinematics remains
lacking. An investigation of the Hercules members' color--magnitude
diagram found no evidence for any significant metallicity anomaly with
respect to other local stars \citep{Bovy10a}, even though such a
difference should be expected as the Hercules stars originate from a
few kiloparsec toward the Galactic center where the average
metallicity is higher than in the Solar neighborhood. The absence of
the predicted metallicity anomaly could potentially be explained
through, \eg, radial mixing \citep{sellwood02a}, but it is clear that
chemodynamical modeling of the Galactic disk is currently not up to
the task of separating out these effects.

@@

To simulate the effect of the bar on stellar orbits near the Solar
circle, we follow the approach of \citet{dehnen00a}. For a fixed
position in the Galaxy, we evaluate the velocity distribution function
at values for the radial---toward the Galactic center---and
tangential---in the direction of Galactic rotation---velocity by
backward-integrating these velocities in the Galactic potential to
obtain the intial orbit before bar formation. We posit that at this
epoch the old stellar disk was in a steady-state that can be described
by a simple distribution function that is a function of the energy and
angular momentum of the orbit alone. It follows from the collisionless
Boltzmann equation that the value of the velocity distribution at the
present position and velocity is equal to that of the steady-state
distribution function before bar formation evaluated at the initial
orbit.

Since the Hercules stream is only apparent in the planar motions in
the disk \citep[\eg,][]{Bovy09a}, we only consider a two-dimensional
model for the Galaxy. We use a simple power-law rotation curve to
model the axisymmetric Galactic potential
\begin{equation}
v_c(R) = \vo(R/\Ro)^\beta \,,
\end{equation}
and for most of the simulations below we specify this model further to
a flat rotation curve ($\beta = 0$). The model for the bar is given by
the following potential
\begin{equation}
\Phi_b(R,\phi) = \Ab(t) \cos [2(\phi - \Omegab t)] \times
\left\{ \begin{array}{ll} -(\Rb/R)^3\,, & \mathrm{for}\ R \geq
  \Rb\,,\\ (R/\Rb)^3-2\,, & \mathrm{for}\ R \leq \Rb\,. \end{array}
\right.
\end{equation}
Here \Omegab\ is the pattern speed of the bar and \Rb\ is the bar
radius, which is fixed to be 80\,percent of the bar's corotation
radius. The bar is grown smoothly during a time $t_1$, always set to
half of the total integration time in what follows, following the
prescription
\begin{equation}
\Ab(t) = \alpha \, \frac{\vo^2}{3}\,\left(\frac{\Ro}{\Rb}\right)^3
\times \left\{ \begin{array}{lll}
\frac{3}{16}\xi^5 - \frac{5}{8} \xi^3 + \frac{15}{16} \xi + \frac{1}{2}\,, & \xi \equiv 2\frac{t}{t_1}-1\,, & t \leq t_1\,,\\
1\,, & & t \geq t_1\,.\end{array} \right. 
\end{equation}
A comparison of the local velocity distribution obtained by using this
simple bar potential with that resulting from the more realistic bar
potential of \citet{gardner10a} shows that this simple model is
sufficient to model the bar's influence on stellar orbits in the outer
disk.

For the distribution function of the stellar disk before bar
formation, we use a Dehnen distribution function \citep{dehnen99b}
given by
\begin{equation}\label{eq:fdehnen}
\fdehnen(E,L) \propto \frac{\Sigma(\rE)}{\sigmaR^2(\rE)} \, \exp\left[ \frac{\Omega(\rE)\left[L-\Lc(E)\right]}{\sigmaR^2(\rE)}\right]\,,
\end{equation}
where \rE, \Lc, and $\Omega(\rE)$ are the radius, angular momentum,
and angular frequency, respectively, of the circular orbit with energy
$E$. Using the procedure given in \sectionname~3.2 of
\citet{dehnen99b}, we choose the $\Sigma(R)$ and $\sigmaR(R)$
functions such that they reproduce a disk with exponential surface
density and velocity dispersion profiles
\begin{equation}
\Sigma(R) = \Sigma_0 \exp\left(-R/\Rs\right)\,, \qquad 
\sigmaR(R) = \sigma_0 \exp\left(-R/\Rsigma\right)\,,
\end{equation}
to an accuracy of a fraction of a percent at all radii. Our fiducial
model for the distribution function has \Rs = \Ro /3 and \Rsigma =
3\Rs, and $\sigma_0$ such that $\sigmaR(\Ro) = 0.2 \vo$.

As shown by \citet{dehnen99c,dehnen00a}, the model described above
with the Sun at \Ro = 0.9\Rolr---where \Rolr\ is the radius at which a
circular orbit is in the outer Lindblad resonance with the bar---at a
current angle of 25$^{\circ}$ with the bar, with a bar strength
$\alpha = 0.01$, and integrating for four bar periods, reproduces the
Hercules feature in the local velocity distribution while agreeing
with other photometric and kinematical observations of the bar
\citep[\eg,][]{binney97a,bissantz02a,cole02a,Shen10a}. In what
follows, unless indicated otherwise, we fix all of the parameters of
the model at these fiducial values. All velocities are always
considered to be with respect to their local standard of rest in the
absence of a bar.

@@

\citet{dehnen00a} only simulated the velocity distribution at the
position of the Sun for various combinations of the model parameters
defined in the previous Section. Nevertheless, his results on how the
velocity distribution at the Sun changes with the angle between the
Sun-Galactic center line and the bar could also be interpreted as
showing the velocity distribution on the Solar circle for different
Galactocentric azimuths, fixing the bar angle. In this way, these
simulations become predictions for what one might find when one
studies the velocity distribution in regions beyond those probed by
the \hipparcos\ mission.

In \figurename~\ref{fig:rphi2d}, I present the predictions for the
velocity distributions near the Solar circle for the full
360$^{\circ}$ range in Galactocentric azimuth for the fiducial model
from \sectionname~\ref{sec:method}. Only the range $-90^{\circ} \leq
\phi \leq 90^{\circ}$ is shown here, since the barred Galaxy is
symmetric with respect to rotations through 180$^{\circ}$, such that
the predictions for the other side of the Galaxy can be obtained from
those shown by shifting the azimuths by 180$^{\circ}$. The $R = \Ro$
locus is not shown here, but it can be inferred from \figurename~2 in
\citet{dehnen00a}.

It is clear from this \figurename\ that we can trace the Hercules
stream going around the Galaxy and when looking further in toward the
Galactic center or further out. The effect of the bar is also not
confined to a narrow range in radii, rather the effects of the bar and
especially its outer Lindblad resonance are felt through a large range
of radii (0.6 $\leq R/\Ro \leq 1.4$). These effects are strong and
robust predictions of the bar-origin model for the Hercules stream,
especially in regions beyond the Solar radius, where the Hercules
feature can be found significantly removed from the bulk of the
velocity distribution, which might be severely shaped at all locations
by the processes that create the pattern of low-velocity moving groups
in the local velocity distribution (the main mode in
\figurename~\ref{fig:obs}). As I discuss below, \Gaia\ will be able to
verify these predictions.

@@

Rather than wait for a comprehensive survey of the Galactic disk that
measures the full phase-space position of a large number of stars by
measuring proper motions, we can ask whether a targeted survey of
line-of-sight velocities in selected regions of the Galaxy could see a
clear signature of the Hercules feature. That this should be possible
follows from the following argument. The Hercules feature stands out
in the velocity distribution most in its tangential velocity
$\vphi$. On the Solar circle near the Sun we see almost entirely this
Galactocentric tangential velocity in the line-of-sight, so we would
expect to see at least a hint of the Hercules stream in the line-of
sight.

In \figurename~\ref{fig:rphi1d} I show the predicted line-of-sight
velocity distributions in various locations near the Solar circle all
around the Galaxy, obtained by integrating the full two-dimensional
velocity distribution over the tangential velocity. This
\figurename\ is no longer symmetric with respect to shifts of
180$^{\circ}$ degree in azimuth, since the line-of-sight direction
breaks this symmetry. The gray curve is the line-of-sight distribution
one would observe if the targeted stars were in a steady state in an
axisymmetric potential with a Dehnen distribution function
(\eqnname~[\ref{eq:fdehnen}]). Both line-of-sight velocity
distributions are normalized to have equal mass since I assume that
our knowledge of the Galaxy and the tracer population will not be
great enough that we can test predicted differences in the absolute
number of stars at a given location between the axisymmetric and the
bar-resonance predictions. These differences are also largely a
function of the particular form and parameters of the distribution
function that we choose the represent the stellar disk before bar
formation.

We see that the prospects to see the Hercules feature are quite good
on our side of the Galaxy. The Hercules feature is especially strong
in the line-of-sight velocity distribution for Galactic longitudes $l
\geq 250^{\circ}$. This is because of the fortunate coincidence that
for those lines-of-sight we see velocities in the direction along
which the Hercules streams separates most from the bulk of the
velocity distribution (see \figurename~\ref{fig:rphi2d}) and that at
these azimuths the Hercules stream is at the furthest distance from
the rest of the velocity distribution. We see hardly any effect of the
bar on the line-of-sight velocity distributions on the other side of
the Galaxy ($90^{\circ} \leq \phi \leq 270^{\circ}$). This is due to
the fact that on that side we see mostly the Galactocentric radial
velocities, in which the Hercules feature does not separate clearly
from the bulk of the stars. Even though upcoming surveys such as
\apogee\ have the ability to probe the other side of the Galaxy, for
the purposes of detecting the Hercules stream this is uninteresting.

To further analyze the best directions to target to best see the
Hercules feature in the line-of-sight velocities, I have calculated
the line-of-sight velocity distribution on a 100 by 20 grid in
Galactocentric azimuth and radius. As a measure of the difference
between the bar-resonance model and the steady-state, axisymmetric
model, I have calculated the Kullback-Leibler divergence between these
predicted distributions \citep[\eg,][]{mackay}
\begin{equation}
D_{\text{KL}}(P||Q) = \int \dd x\,p(x)\,\ln \frac{p(x)}{q(x)}\,,
\end{equation}
where, in this case, $x$ is the line-of-sight velocity, $p(x)$ is the
predicted distribution for the bar-origin model of the Hercules
stream, and $q(x)$ is the axisymmetric prediction. The
Kullback-Leibler divergence is only one among many possible choices to
quantify the difference between two probability distributions. It can
be interpreted as the expected weight of evidence for the prediction
by the bar-resonance model over the fully axisymmetric prediction that
can be expected from each sampled star. As such, locations with the
maximal Kullback-Leibler divergence are those in which the Hercules
feature stands out most prominently in the line-of-sight.

@@

Since the extinction in the optical bandpasses is rather large in the
directions identified as promising in \figurename~\ref{fig:detect}
\citep[$A_V \approx 8$ mag toward $l \approx
  270^{\circ}$;][]{marshall06a}, spectroscopic surveys in the
near-infrared such as \apogee\ are the best bet for seeing the
Hercules feature in the line-of-sight direction. The range of Galactic
longitudes that \apogee\ can target along the disk is $-5^{\circ} \leq
l \leq 250^{\circ}$ (SDSS-III Project Description\footnote{Available
  at \url{http://sdss3.org/collaboration/description.pdf}~.}). From
\figurename~\ref{fig:detect} this limit seems to be right on the edge
of where one might hope to detect the Hercules feature. Since
\apogee\ is expected to start taking data in 2011 and to finish
operations before the first \Gaia\ data release, \apogee\ could trace
the Hercules stream around the Galaxy before \Gaia\ does. Since
\apogee\ will also obtain detailed chemical abundances for its target
stars, this would also allow for a more detailed analysis of the
abundances of stars in the Hercules stream \citep[following,
  \eg,][]{Bensby07a,Bovy10a}, which could shed more light on radial
mixing in the Galactic disk.

@@

\bibitem[Caloi \etal(1999)]{caloi99a}
  Caloi,~V., \etal, 1999,
  \aap, 351, 925

@@

\bibitem[Marshall \etal(2006)]{marshall06a}
  Marshall,~D.~J., \etal, 2006,
  \aap, 453, 635

@@

\clearpage
\begin{figure}
\includegraphics[width=0.5\textwidth]{veldist.eps}
\caption{The velocity distribution of nearby ($\lesssim 100$ pc) stars
  from \citet{Bovy09a}. \vR\ and \vphi\ are the velocities toward the
  Galactic center and in the direction of Galactic rotation,
  respectively, corrected for the Solar motion. The Hercules moving
  group is visible as an overdensity at \vphi $\approx$ -0.2
  $\vo$.}\label{fig:obs}
\end{figure}

\clearpage
\begin{figure}
\includegraphics[width=\textwidth]{rphi2d.ps}
\caption{Two-dimensional velocity distributions along the Solar
  circle. The axes of each subpanel are the same as in
  \figurename~\ref{fig:obs}. The velocity distributions on the other
  side of the Galaxy are obtained by shifting the azimuths by
  180$^{\circ}$.}\label{fig:rphi2d}
\end{figure}

@@

\clearpage
\begin{figure}
\includegraphics[width=0.5\textwidth]{distuncertain.ps}
\includegraphics[width=0.5\textwidth]{df.ps}\\
\includegraphics[width=0.5\textwidth]{slope.ps}
\includegraphics[width=0.5\textwidth]{barstrength.ps}
\caption{Variation of the predicted velocity distribution at R = 1.075
  \Ro , $\phi = -21^{\circ}$ (d= 0.4 \Ro, $l$ = 270$^{\circ}$) with
  distance uncertainties, parameters and shape of the distribution
  function, shape of the rotation curve, and bar
  strength.}\label{fig:1dvar}
\end{figure}
